\documentstyle[A4]{article}

\input amssym.def
\newcommand{\Rr}{\Bbb R}
\newcommand{\Qq}{\Bbb Q}

\title{Statement of purpose}

\author{Jo\~ao Paulo Neves Monteiro dos Santos}

\begin{document}

\maketitle

I've chosen to take an undergraduate degree in physics
because of my desire to under\-stand the foundations of science.
While taking this degree I realized that physics was not what
I have thought it was. What I wanted was not to discover the theories
that better
explain the observed phenomena but an understanding of the theories
themselves.
Most physical theories are not really well understood.
As a mathematician I want 
to formalize the mathematics they use and seek a better understanding of the 
organization of the world.

During my undergraduate studies, I have followed
several advanced courses in mathematics because I felt that what I knew
was not enough to understand the physical
theories but also because mathematics always interested me by itself.
Through them I became more and more interested in analysis.
Presently I'm particularly interested in those
fine questions in analysis
related to the characterization of properties of solutions of
differential equations, problems of the calculus of variations, geometric
measure theory and their consequences for physical theories. However, since I'm
not acquainted with many of other topics I'm not sure what I'll choose at the
end.

I want to pursue my studies essentially because
I want to learn new things and I want to be in a challenging environment in
top research, where I can contact a varied group of active mathematicians.
So I'm interested in a school with regular research seminars and
a strong visitor's program. Anyway,
before I choose a specific Ph.D. thesis topic I want to acquire a strong
background in several different fields of mathematics
because I would like to stand on deep and solid bases.
Following that line  of thought I had decided to do my master thesis on
Integrable Hamiltonian Systems since to that I had to study Lie Theory and
Differential Geometry, something I knew nothing about just a few months ago. One
important reason why I've chosen to get a Ph.D. in the United States is that its
graduate programs are well structured and have larger and much more organized
course components than most of the places in Europe. That allows for a period of
acquisition of wider and deeper knowledge before the attack of specific research
problems.

For the past $2{1\over 2}$ years I have been teaching advanced
calculus to first and second year undergraduate students in
mathematics, physics and  engineering. This experience has given me  a better
understanding of some very basic but very important concepts and I
could confirm in practice the impression that I would
like teaching. I want to follow a career in research and teaching
at the university because in doing so I can conciliate my desire to learn and
find out new things with my taste for teaching.

\end{document}